Stack Safety is a Security Property

### 2021-04-17 tex/main.tex
Finished pass

We propose a formal characterization
of stack safety, formulated with concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We use these properties to validate the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to a slightly weaker and more extensional
variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

The call stack is a perennial target for low-level attacks, leading to
consequences ranging from leakage or corruption of private stack data to
control-flow hijacking. To prevent or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate,
and they often require trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{SkorstengaardSTK}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  Our technical formulation of WBCF is similar to
that of Skorstengaard et al.{}, but our treatment of stack data protection
is rather different.  Instead of defining it implicitly,
as ``what a certain idealized machine does,'' we give an
\emph{explicit} definition using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi, phrased as the conjunction of
\emph{stack integrity} and \emph{stack confidentiality}.
%
A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:control}).
\rb{\cref{sec:RTC}?}
This reflects an arguably cleaner division between data protection and
control-flow protection than LSE / WBCF.

For the confidentiality and integrity properties, we consider both {\em
  stepwise} and {\em observational} variants.  The simpler {stepwise}
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The more extensional observational variants
allow a callee to read from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.

@@

We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but
testing
reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
stepwise or
observational): it violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee.  We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
confidentiality (as far as our testing can tell\bcp{right, leo?}), albeit at
some performance cost.
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' macro-instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

Returns, on the other hand, need not be annotated: we can tell when a callee
has returned control to the caller when... we see control return to the
caller!  For example, under a standard calling convention, we can see that a
return has occurred occurred when state of the machine reaches a state where
the stack pointer has the same value as it did at the point of some earlier
call and the $\PCname$ points to the instruction following that call.  We
abstract away from the details of specific calling conventions by assuming
that we are given some way of telling when a given machine state is a
{\em return target}\bcp{with respect to some set of pending calls?}.

@@

Intuitively, stack safety means that the stack frame of each caller (and
those of its ancestor callers) is protected
from attempts by a confused or malicious callee (including all its
descendant callees)
to read or write its private data.
That is, the attacker model treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function invocation
may be viewed in both ways.

@@

% \paragraph*{Well-formedness and Bad Annotations}
% \bcp{What is ``well-formedness''??}

Since we rely on call annotations\bcp{to...}, we must ask: what if the annotations
are wrong? First, suppose a call occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call, only an instruction sequence resembling one, and
therefore no associated change in permissions. Then the caller is guaranteed no
protections.

\bcp{Confusing, and maybe out of date?}Alternatively, an annotated instruction might not obey the calling convention.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish, termed {\em blessed sequences}.
When applying our properties to a micro-policy setting, we will statically rule out some
scenarios with a well-formedness criterion on initial states: all annotated instructions will
be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.

@@

10 f:   add $1,%sp      ; allocate frame
11      mov $0,-3(%sp)  ; violation!
12      mov $5, %ra     ; set return value
13      sub $1,%sp      ; deallocate frame
14      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}[t]{.25\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 2} &
\memoryaddrs{8em}
\memory{3}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 3} &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 12} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}
\vspace{-4ex}
\end{subfigure}
\caption{An integrity violation
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

@@

Colored boxes indicate an assignment of the address to a
{\em domain}---a label reflecting a particular set of access rights. Domains
are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
variables. Domains are part of the {\em context} of a state.
We will introduce the remaining context as needed.

@@

\paragraph{Well-bracketed control flow.} Together,
integrity and confidentiality protect a caller's data until its callee
returns to it. However, it is possible that an intermediate caller is not
returned {\em to}, but rather returned {\em past}, to some ancestor caller!
For example, function $A$ calls function $B$, which in turn calls $C$, which
returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
and branching to the address after the call from $A$ to $B$.  Strictly
speaking, this is not an information-flow violation, since neither $A$'s nor
$B$'s stack frames are improperly read or written by $C$.  But it is still
arguably unsafe in general.  For example, suppose that $B$'s continuation
after the call to $C$ takes $C$'s result, doubles it, and returns this to
$A$.  If $A$'s correct behavior after the call to $B$ returns depends on the
result it receives being a even number---i.e., $A$ is trusting that $B$'s
continuation will execute after $C$ returns, perhaps because $A$ and $B$ are
written by the same person while $C$ is an untrusted library---then $C$ can
cause it to behave incorrectly, possibly even insecurely, by returning an
odd result directly to $A$.
%
Thus, a formal definition of safe use of the stack should include the
condition that calls and returns are {\em well bracketed}, e.g., $C$ should
only return to $B$.  This is what \citet{SkorstengaardSTK} describe
as \emph{well-bracketed control flow (WBCF)}.

@@

We are finally ready for our definition of stack safety. We begin
by describing how we model knowledge about the program structure,
then dive into the formal definitions of local state encapsulation\apt{we don't want
  to call it that here, do we?} and
well-bracketed control flow, which together constitute stack safety.

We will focus on a simple stack model, with a single stack
that doesn't share between caller and callee: arguments and results are passed
in registers. To accommodate different machine architectures and different calling
conventions, we parameterize over the structure of the stack, particularly
the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention}, \(\sealcon \subseteq \MACHS \times \COMPONENTS\), a
relation between machine states and the addresses that should be sealed.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return. Similarly, we abstract over what
it means to return with a {\em return convention}, \(\retcon \subseteq \MACHS \times \MACHS\),
which relates call states to the states that can be considered valid returns to that call.
The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).

Each system component is assigned to a {\em domain}, which may be {\em Outside}
of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
at a particular depth. Each caller registers a {\em target}
condition---a predicate on machine states that must hold for it to consider itself active.
A context is a pair of a ``domain map'' from components to domains, and a
``return target list,'' which is a stack of targets, one for each caller awaiting
return.  \ifaftersubmission\bcp{spacing is funny}\fi
%
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
  \times (\listT (\target)) \]
%
The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
\rb{Technically a partial function from addresses to annotations,
  though its domain can function as a predicate.}

@@

\definition
A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and any component \(\component \in \prot(\domMap,\rts)\),
\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
\apt{but now $dm$ and $rts$ are unbound}.

@@

\definition
For any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
let \(\components = \mathit{secret}(\context)\).
A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
for any \(\nach\) such that \(\mach \approx_\components \nach\) and the induced trace
\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
  universally quantified, and probably also \(\MPCT\) (?) }
\sna{Better?\apt{only partly. Quantification of $M,\brk$ still seems wrong to me. }}

@@

We extend this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we sometimes allow the generation to relax those constraints,
introducing potentially ill-formed flows: this causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.

>>>>

We propose a formal characterization
of stack safety, formulated with concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We use these properties to validate the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to a slightly weaker and more extensional
``observational'' variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

The call stack is a perennial target for low-level attacks, leading to
consequences ranging from leakage or corruption of private stack data to
control-flow hijacking. To prevent or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate,
and they often require trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees these properties, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also describe the protection of both stack data (like LSE) and
control flow (like WBCF).  Our technical formulation of WBCF is similar to
theirs, but our treatment of stack data protection is rather different,
as is our approach to the definition itself. Instead of defining stack safety implicitly,
as ``what a certain idealized machine does,'' we give an
\emph{explicit} definition using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi, phrased as the conjunction of
\emph{stack integrity} and \emph{stack confidentiality}.
%
A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of WBCF, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
This reflects an arguably cleaner division between data protection and
control-flow protection than LSE / WBCF.

For confidentiality and integrity , we consider both {\em stepwise}
and {\em observational} variants.  The simpler stepwise
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The observational variants
allow a callee to read from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.
The observational properties are meant to capture the same sense of security,
extensionally enough that any reasonable protection mechanism ought to enforce
them, even if it does not prevent every single dangerous read or write.

@@

We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties.  On the other hand, testing reveals their more efficient
\emph{Lazy Tagging} policy does not enforce step-wise integrity. In fact
we reason that both of their lazy policies violate the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and \emph{Lazy Tagging and Clearing} also violates integrity in these instances. We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
confidentiality, albeit at some performance cost.
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy \sna{RISCy?} ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' macro-instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

Returns, on the other hand, need not be annotated: we can tell when a callee
has returned control to the caller when... we see control return to the
caller!  For example, under a standard calling convention, we can see that a
return has occurred occurred when state of the machine reaches a state where
the stack pointer has the same value as it did at the point of some earlier
call and the $\PCname$ points to the instruction following that call.  We
abstract away from the details of specific calling conventions by assuming
that we are given some way of telling when a given machine state is a
{\em return target} with respect to some set of pending calls.

@@

Intuitively, stack safety means that the stack frame of each caller is protected
from attempts by a confused or malicious callee (including all its
descendant callees) to read or write its private data.
That is, the attacker model treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function invocation
may be viewed in both ways. We do rule out some
scenarios with a ``well-formedness'' criterion on initial states, ensuring that
they are set up properly for the enforcement mechanism in question.
When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
well-behaved calls and returns are implemented by fixed sequences of instructions
which the tag policy will force to be executed from start to finish,
termed {\em blessed sequences}. Well-formedness requires that all annotated
instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.

@@

@@

10 f:   add $1,%sp      ; allocate frame
11      mov $0,-3(%sp)  ; violation!
12      mov $5, %ra     ; set return value
13      sub $1,%sp      ; deallocate frame
14      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}[t]{.25\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 2} &
\memoryaddrs{8em}
\memory{3}{\unsealc}[U]
~$\cdots$
\vspace{.5em}
\\
{\tt 3} &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[S(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[U]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 12} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}
\vspace{-4ex}
\end{subfigure}
\caption{An integrity violation
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

@@

Colored boxes indicate an assignment of the address to a
{\em domain}---a label reflecting a particular set of access rights. Domains
are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
figures. For our purposes, \(\outside\) will apply to global
variables. Domains are part of the {\em context} of a state.
We will introduce the remaining context as needed.

@@

\paragraph{Well-bracketed control flow.} Together,
integrity and confidentiality protect a caller's data until its callee
returns to it. However, it is possible that an intermediate caller is not
returned {\em to}, but rather returned {\em past}, to some ancestor caller!
For example, function $A$ calls function $B$, which in turn calls $C$, which
returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
and branching to the address after the call from $A$ to $B$.  Strictly
speaking, this is not an information-flow violation, since neither $A$'s nor
$B$'s stack frames are improperly read or written by $C$.  But it is still
arguably unsafe in general.  For example, suppose that $B$'s continuation
after the call to $C$ takes $C$'s result, doubles it, and returns this to
$A$.  If $A$'s correct behavior after the call to $B$ returns depends on the
result it receives being a even number---i.e., $A$ is trusting that $B$'s
continuation will execute after $C$ returns, perhaps because $A$ and $B$ are
written by the same person while $C$ is an untrusted library---then $C$ can
cause it to behave incorrectly, possibly even insecurely, by returning an
odd result directly to $A$.
%
Thus, a formal definition of safe use of the stack should include the
condition that calls and returns are {\em well bracketed}, e.g., $C$ should
only return to $B$.  This is what \citet{Skorstengaard+19b} describe
as \emph{well-bracketed control flow (WBCF)}.

@@

We are finally ready for our definition of stack safety. We begin
by describing how we model knowledge about the program structure,
then dive into the formal definitions of integrity, confidentiality, and
well-bracketed control flow, which together constitute stack safety.

We will focus on a simple stack model, with a single stack
that doesn't share between caller and callee: arguments and results are passed
in registers. To accommodate different machine architectures and different calling
conventions, we parameterize over the structure of the stack, particularly
the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention},
\(\sealcon ~ (\mach \in \MACHS) ~ (\component \in \COMPONENTS)\), a
relation between machine states and the addresses that should be sealed.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return. Similarly, we abstract over what
it means to return with a {\em return convention},
\(\retcon ~ (\mach \in \MACHS) ~ (\mach' \in \MACHS)\),
which relates call states to the states that can be considered valid returns to that call.
The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).

Each system component is assigned to a {\em domain}, which may be {\em Outside}
of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
at a particular depth. Each caller registers a {\em target}
condition, \(\target ~ \mach\)---a predicate on machine states that means it
has been returned to.
A context is a pair of a ``domain map'' from components to domains, and a
``return target list,'' which is a stack of targets, one for each caller awaiting
return.
%
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} \hspace*{0.5in} \target \subseteq \MACHS\]
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
  \times (\listT (\target)) \]
%
The initial context \(\context_0 = \domMap_0, []\), where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is a predicate on addresses.

@@

\definition
A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
\(\mach,\pol,\context\) such that \(\mach,\pol,\context \stepstocon \mpcstatename'\),
and any component \(\component \in \prot(\context)\),
\(\pi_\mach(\mpcstatename') ~ \component = \mach ~ \component\).

@@

\definition
A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
for any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
\(\nach\) such that \(\mach \approx_{\mathit{secret}(\context)} \nach\) and the induced trace
\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
\(\MPCT \doteq \NPCT\) holds.

@@

We extended this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we sometimes allow the generation to relax those constraints,
introducing potentially ill-formed flows: this causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.
